El Mayor Postseismic \author{Trever Hines} \rule{\headwidth}{1.0pt}

### 2016-03-29 RoughDraft/ElMayor.tex
commit before push

\usepackage{amssymb,amsmath}
\usepackage[margin=1.0in]{geometry}
\usepackage{fancyhdr} % required for custom header
\usepackage{graphicx}
\usepackage{listings}
\usepackage{courier}
\usepackage[usenames,dvipsnames]{color}
\usepackage[maxfloats=30]{morefloats}
\usepackage{caption}
\usepackage{subcaption}

@@

Previous studies which have modeled postseismic deformation following the El Mayor-Cucapah earthquake include \cite{Pollitz2012}, \cite{Gonzalez-ortega2014}, \cite{Spinler2015}, and \cite{Rollins2015}. \cite{Gonzalez-ortega2014} was able to describe five months of near field ($\lesssim 50$ km from the epicenter) postseismic defomation observed by InSAR and campaign GPS with afterslip and fault contraction on the coseismically ruptured fault. \cite{Gonzalez-ortega2014} also noted that their preferred model underestimated the GPS displacements for stations $\gtrsim 25$ km from the rupture and suggested that it could be the result of unmodeled viscoelastic relaxation.  Using continuous GPS stations within 200 km of the El Mayor-Cucapah epicent, which consists mostly of stations north of the US-Mexico border, \cite{Rollins2015} found that three years of postseismic deformation can be adequately explained by afterslip in an elastic lithosphere, albeit with an implausibly large amount of afterslip inferred on the least constrained southern most fault segment which may be acting as a proxy for distributed relaxation in the lower crust or upper mantle. In this paper, we reiterate this finding by \cite{Rollins2015} and note that a purely elastic dislocation model is incapable of describing deformation observed at GPS stations greater than ~200 km from the El Mayor-Cucapah epicenter.

Given the inability to describe both near and far field deformation with fault slip in an elastic lithosphere, \cite{Pollitz2012}, \cite{Rollins2015} and \cite{Spinler2015} have explored viscoelastic relaxation in the lower crust and upper mantle as a potential deformation mechanism. The lithospheric rheology is largely unknown and so modeling postseismic deformation with viscoelastic relaxation requires one to assume a rheologic model and find the best fitting model parameters, which is generally a computationally expensive nonlinear inverse problem. Consequently, a simplified structure for the lithosphere must be made to minimize the number of rheologic parameters that need to be estimated.  For example, it is commonly assumed that the lithosphere consists of only three homogeneously Maxwell viscoelastic layers, which may be an inadequate representation of the lithosphere \cite{Hines2013}\cite{Riva2009}. Additionally, it is necessary to make simplifying assumptions about the nature of afterslip. For example, one can assume a frictional model for afterslip and parameterize afterslip in terms its the unkown rheologic properties (e.g. \cite{Johnson2009} \cite{Johnson2004}). Additionally, one can assume that afterslip only dominates the first year of deformation (e.g. \cite{Pollitz2012} and \cite{Spinler2015}) and that viscoelastic relaxation is the dominant mechanism in later years. However, numerous observations of interseismic fault creep has been recognized in this and similar tectonic settings and it has been speculated that such creep can be initiated as a postseismic process \cite{Cakir2012} \cite{Cetin2014}.  We are unaware of any oIn this study, we therefore do not make to implicit assumption that afterslip terminates after a given amount of time.  Indeed, the preferred viscoelastic model from \cite{Pollitz2012} underpredicts near field velocities, which could be indicative of unmodeled continued afterslip.

All of the aforementioned studies model displacements observered at stations within 200 km of the El Mayor-Cucapah epicenter, while postseismic deformation in this region following previous earthquakes of similar magnitude has been observed at distances extended out to about 300 km \cite{Freed2007a}. In the present study we examine stations within 400 km of the El Mayor-Cucapah epicenter, which turns out to be crucial for discerning the postseismic deformation mechanism.

Clearly, both afterslip and viscoelastic relaxation are involved in postseismic deformation neglecting to model one of these deformation mechanisms could result in a biased estimate of the other.  In this paper we use the inverse method described in \cite{Hines2015} to estimate the afterslip an effective viscosity necessary to describe the transience postseismic deformation over the first ten months after the El Mayor-Cucapah earthquake. We then form a suite of models which have various lithospheric rheologies but viscosities that are consistent with the effective viscosity estimated from the early postseismic deformation. Of the suite of models tested, we find that five years of postseismic deformation can be explained by a combination of sustained afterslip on the coseismicly ruptured fault and a Zener rheology upper mantle with viscosity that decays from $4\times10^{18}$ to $1\times10^{18}$ Pa s.

@@

We use continuous GPS position time series provided by University Navstar Consortium (UNAVCO) for Plate Boundary Observatory (PBO) stations within a 400 km radius about the El Mayor-Cucapah epicenter. Our analysis is on the coseismic and postseismic deformation resulting from the EMC earthquake, which we collectively describe as $u(t)$. We consider GPS position time series, $u_\mathrm{obs}(t)$, to be the superposition of $u(t)$, secular tectonic deformation, annual and semi-annual fluctuations, and coseismic offsets from sigificant earthquakes over the time span of this study.  The June 14, 2010 Mw5.8 Ocotillo earthquake and the August 26, 2012 Brawley swarm, which consisted of a Mw5.5 and Mw5.3 event, are the only earthquakes after the EMC earthquake that produced noticeable offsets recorded by GPS. Although the Ocotillo earthquake had its own series of aftershocks (Haukson), neither earthquake produced transient deformation that is detectable with GPS. We thus model $u_\mathrm{obs}(t)$ as

@@

Stations which recorded signals that clearly cannot be described by the aforementioned processes are not included in our analysis. This includes stations in the Los Angeles basin, which record deformation that is largely anthropogenic. In order to ensure an accurate estimation of the secular deformation, we only use stations that were installed at least six months prior to El Mayor-Cucapah earthquake. While several stations were installed after the EMC earthquake to improve the spatial resolution of postseismic deformation \cite{Spinler2015}, our inverse method uses postseismic displacements rather than velocities (e.g. ...)
ELABORATE ON VELOCITY VS DISPLACEMENT
, which requires the knowledge of the stations preseismic position. Despite our inability to utilize potentially rich data, we prefer to use postseismic displacements rather than potentially dubious estimates of postseismic velocities.

The October 16, 1999 Hector Mine earthquake, which occurred within our study region about 270 km north of the EMC epicenter, has produce transient postseismic deformation which we do not wish to model either mechanically or through empirical line fitting. We thus restrict our analysis to deformation observed six years after the Hector Mine earthquake, past which point postseismic deformation for nearfield sites occurs at an approximately steady rate \cite{Savage2009}. When considering stations further away from the Hector Mine epicenter, postseismic transience persists for only about two years \cite{Spinler2015}.

Studies of postseismic deformation typically assume a parametric form for $u(t)$ such as such as a logarithmic or exponential function. Such functions have been found empirically e.g. \cite{Savage2005a} and theoretically \cite{Marone1991} to adequately describe deformation resulting from postseismic deformation. However, by assuming a logarithmic or exponential form of $u(t)$ we run the risk of overfitting the GPS time series and inferring a nonexistent postseismic signal. We therefore do not assume any parameteric form for $u(t)$ and rather treat it as integrated Brownian motion, so that 
\begin{equation}
    \dot{u}(t) = \sigma^2\int_0^t w(s) ds.
\end{equation}

@@

The smoothness of $u(t)$ is controlled by the chosen value of $\sigma^2$, which describes how rapidly we expect the postseismic signal to vary over time.  Setting $\sigma^2$ equal to zero will effectively result in modeling $u(t)$ as a straight line which is insufficient to describe the expected transient behavior in postseismic deformation \cite{Savage2005a}. The other endmember, where $\sigma^2$ is infinitely large, will result in $u_\mathrm{pred}(t)$ fitting what is obviously noise in the data. While one can use a maximum likelihood based approach to picking $\sigma^2$ (e.g. \cite{Segall1997}), we rather take a subjective approach and choose a value for $\sigma^2$, which is used when filtering time series for each station, that is just large enough to faithfully describes the early rapid rate of postseismic deformation at the most near field station in our study, P496.  This ensures that $\sigma^2$ will be sufficiently large so that our estimate of $u(t)$ does not smooth out potentially valuable postseismic signal. We find that when using $\sigma^2 = 0.05 \mathrm{m}^2 / \mathrm{yr}^3$, we are able to adequately describe all but the first week of postseismic deformation at station P496, which gets incorporated into our estimate of coseismic displacements (figure \ref{P496Fit}).  Since we assume that the first week of deformation is overwhelmingly the result of afterslip, any unmodeled afterslip over the first week following the El Mayor-Cucapah earthquake will be added into our estimates of coseismic slip in section ().  Figure \ref{P496PS} shows the estimate of coseismic and postseismic deformation, $u(t)$, for station P496 along with its formal uncertainties.

It is important to note that the shown uncertainties in $u(t)$ do not account for the non-negligible epistemic uncertainty in eq. \ref{TimeSeriesModel}.  For example, we assume that the background rate of deformation can be approximated as having a constant rate, while mechanical models would suggest this to not be true on earthquake cycle timescales \cite{Thatcher1983}. Although based on inspection,  assuming a constant rate of background deformation appears to be appropriate for all stations except for perhaps the stations closest to the Hector Mine epicenter, where postseismic deformation persists.  Also, our model for seasonal deformation in eq. \ref{TimeSeriesModel} assumes a constant amplitude over time, which means that years of particularly heavy or light rainfall cannot be adequately described.  This deficiency in our seasonal model causes our estimate of $u(t)$ to describe some of the unmodeled oscillations (figure \ref{P619PS}).

@@

Another reason for constraining slip to be shallower than 15 km depth is because we seek to estimate a lower crustal viscosity and the inverse problem becomes particularly ill-posed when trying to simultaneously estimate afterslip and a lithospheric viscosity at the same depths. 
This ill-posedness is illustrated in figure (TBD), which shows the displacement resulting from a meter of slip on a fault extending from 15 to 30 km depth as well as the initial velocity resulting from viscoelastic relaxation in the lower crust, which is given a viscosity of $10^{18}$ Pa s.  The horizontal displacement from fault slip is in the opposite direction as the displacements resulting from subsequent viscoelastic relaxation.  This means that displacement resulting from afterslip at lower crustal depths can be cancelled out by a low viscosity lower crust.  We eliminate this null space by allowing only one mechanism in the lower crust, which we choose to be viscoelastic relaxation.  Aside from the issue of nonuniqueness, it is also nonsensical to allow for both brittle and ductile deformation mechanisms at the same depth.

@@

With afterslip constrained to shallow depths, we must now select a viscoelastic rheology for the lower crust and upper mantle. Maxwell viscoelasticity, which can be schematically illustrated as a spring and dashpot connected in series, is the simplest and commonly used rheologic model for studies on postseismic deformation (e.g. \cite{Nur1974}, \cite{Pollitz2000}, \cite{Hetland2003},\cite{Freed2006a}, \cite{Johnson2009}, \cite{Hearn2009}).  The Maxwell model describes both an instantaneously elastic and fluid-like phase of deformation.   Assuming Maxwell viscoelasticity introduces the additional rheologic parameter, $\eta_M$, which is typically estimated for the lower crust and upper mantle as an inverse problem, where the viscosities are chosen to minimize the difference between the observed postseismic deformation and deformation predicted by a mechanical model. A Maxwell viscoelastic model of the lithosphere would predict deformation that decays steadily over time, which fails to describe the observable deformation which exhibits transient deformation which decays rapidly within the first year of an earthquake followed by a relatively steady rate of deformaiton \cite{Savage1997}.  An additional deficiency in describing postseismic deformation with viscoelastic relaxation in a Maxwell viscoselastic lithosphere is that inferred viscosities tend to be considerably lower than viscosities inferred from geophysical processes that occur over longer timescales. For example, postseismic studies in Southern California following the Landers \cite{Pollitz2000}, Hector Mine \cite{Pollitz2001}, and El Mayor-Cucuapah earthquake \cite{Spinler2015}, \cite{Rollins2015}, have all inferred a Maxwell viscoselastic upper mantle with a viscosities ranging from $10^{17}$ to $10^{18}$ Pa s; however, studies on interseismic deformation \cite{Lundgren2009} and lake loading \cite{Luttrell2007}, which are both processes that occur on the timescale of $~10^2$ years, infer a Maxwell viscoelastic upper mantle with a viscosity on the order of $10^{19}$ Pa s.

We also consider the Zener viscoelastic model whose mechanical analogue consists of a spring connected in series with a Kelvin element. The Zener model is used to describe both an instantaneous and delayed phase of elasticity.  A generalization of the Zener model, consisting of numerous Kelvin elements connected in series, is commonly used to describe the spectrum of frequencies where seismic attenuation is observed \cite{Liu1976}.  The highest viscosity needed to describe attenuation of earth's lowest frequency normal modes is on the order of $10^{16}$ Pa s \cite{Yuen1982} and the largest characteristic relaxation time would be on the order of days, which is too short to be perceptible in the GPS data used in the present study. We therefor ignore the delayed elasticity associated with seismic attenuation models, and consider such models to be effectively instantaneously elastic. Nevertheless, we still explore the possibility of describing postseismic deformation with a Zener model that has relaxation timescales of months to years.

Of course, it has long been recognized that a Zener rheology provides an incomplete descriptions of the asthenosphere, as it does not have the fluid behaviour required for mantle dynamics on geologic time scales. \cite{Yuen1982} proposed a Burgers rheology with a low viscosity ($\approx 10^{16}$ Pa s) in the Kelvin element and high viscosity ($\approx 10^{21}$ Pa s) in the Maxwell element which is capable of describing both seismic attenuation and the fluid-like behaviour required for glacial isostatic rebound and mantle convection. The justification of a Burger's rheology mantle is further supported by laboratory experments on the strength of olivine \cite{Chopra1997}. \cite{Pollitz2003} sought to describe postseismic deformation following Hector Mine with a Burgers rheology mantle and they found a best fitting transient viscosity of $1.6\times10^{17}$ Pa s and steady state viscosity of $4.6\times10^{18}$ Pa s. While the Burgers rheology was introduced as a means of bridging to gap between relaxation observed in long and short term geophysical processes, the inferred steady state viscosity from \cite{Pollitz2003} is still inconsistent with the Maxwell viscosities inferred from \cite{Lundgren2009} and \cite{Luttrell2007}. Additionally, the transient viscosity inferred by \cite{Pollitz2003} is needed to describe the earliest phase of postseismic deformation following the Hector Mine earthquake, while \cite{Pollitz2003} ruled out deep afterslip as an alternative mechanism based on inconsistent vertical deformation, it is still possible to successfully describe all components of early postseismic deformation following the Hector Mine earthquake with afterslip at seismogenic depths \cite{Jacobs2002}. In this paper we seek to constrain the rheologic parameters describing a Burgers upper mantle that are consistent with postseismic deformation and long term geophysical processes, while also accounting  for shallow afterslip, which perhaps biased the solution from \cite{Pollitz2003}.

what this gives us is a prior constraint on lithospheric viscosity initial that is unbiased by neglecting the contribution of afterslip!

It can be shown (\cite{Hines2015}),

other rheologic models in this study      viscoelasticity is We also test rheologic models other than viscoelasticity to describe

There are numerous other parameters which are not known, but will be considered known in this study.

   Inherent in f is also a stress-linear viscoelastic model.   In this study we will be be known while try to consider

   give table of assumed and estimated material properties

   The early near field postseismic vertical deformation does show a quadrant pattern and is also consistent with the coseismic vertical deformation.  Given

that would not becomes there is no longer a coherent far field deformation signal. This is in contrast to near field displacements which persists for the entire 5.0 years of data used in this study. Figure () further illustrates the difference in deformation regimes between near field and far field station.

note that subsidence in the peninsular range can be explained with either a localized weakness or conitnued afterslip at seismogenic depths

We use a fault geometry from \cite{Wei2011}, which was determined using teleseismic, GPS, and InSAR data.

make note of \cite{Yamagiwa2014}

note that creep in Brawley is probably the main source of error

After the El Mayor-Cucapah earthquake, additional GPS stations were installed In Baja California to record postseismic deformation with better spatial coverage.  Two of the stations PTAX and PHJX were installed near the epicenter in the Cucapah Mountains.  We do not include these two stations in our analysis because the geometry of the faults that ruptured during the earthquake is more complicated than our assumed fault geometry \cite{Oskin2012} \cite{Fletcher2014} and the near field stations would be most sensitive to error in our geometry. We also leave of these near field stations in order to avoid any near field processes which we do not consider in this paper \cite{Gonzalez-ortega2014}. Numerous stations exhibit extraneous signals which can be

@@

@@

\begin{figure}
\includegraphics[scale=0.9]{Figures/near_field_record_section}
\centering 
\caption{near field record section}
\label{fig:NearFieldRS}
\end{figure}
\begin{figure}
\includegraphics[scale=0.9]{Figures/far_field_record_section}
\centering 
\caption{far field record section}
\label{fig:FarFieldRS}
\end{figure}        
\begin{figure}

\includegraphics[scale=1.0]{Figures/slip_lcurve}
\centering 
\caption{near field record section}
\label{fig:SlipLCurve}
\end{figure}

>>>>

\usepackage{amssymb,amsmath}
\usepackage[margin=1.0in]{geometry}
\usepackage{fancyhdr} % required for custom header
\usepackage{graphicx}
\usepackage{listings}
\usepackage{courier}
\usepackage[usenames,dvipsnames]{color}
\usepackage[maxfloats=40]{morefloats}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{natbib}

@@

Previous studies which have modeled postseismic deformation following the El Mayor-Cucapah earthquake include \citet{Pollitz2012}, \citet{Gonzalez-ortega2014}, \citet{Spinler2015}, and \citet{Rollins2015}. \citet{Gonzalez-ortega2014} was able to describe five months of near field ($\lesssim 50$ km from the epicenter) postseismic defomation observed by InSAR and campaign GPS with afterslip and fault contraction on the coseismically ruptured fault. \citet{Gonzalez-ortega2014} also noted that their preferred model underestimated the GPS displacements for stations $\gtrsim 25$ km from the rupture and suggested that it could be the result of unmodeled viscoelastic relaxation.  Using continuous GPS stations within 200 km of the El Mayor-Cucapah epicent, which consists mostly of stations north of the US-Mexico border, \citet{Rollins2015} found that three years of postseismic deformation can be adequately explained by afterslip in an elastic lithosphere, albeit with an implausibly large amount of afterslip inferred on the least constrained southern most fault segment which may be acting as a proxy for distributed relaxation in the lower crust or upper mantle. In this paper, we reiterate this finding by \citet{Rollins2015} and note that a purely elastic dislocation model is incapable of describing deformation observed at GPS stations greater than ~200 km from the El Mayor-Cucapah epicenter.

Given the inability to describe both near and far field deformation with fault slip in an elastic lithosphere, \citet{Pollitz2012}, \citet{Rollins2015} and \citet{Spinler2015} have explored viscoelastic relaxation in the lower crust and upper mantle as a potential deformation mechanism. The lithospheric rheology is largely unknown and so modeling postseismic deformation with viscoelastic relaxation requires one to assume a rheologic model and find the best fitting model parameters, which is generally a computationally expensive nonlinear inverse problem. Consequently, a simplified structure for the lithosphere must be made to minimize the number of rheologic parameters that need to be estimated.  For example, it is commonly assumed that the lithosphere consists of only three homogeneously Maxwell viscoelastic layers, which may be an inadequate representation of the lithosphere \citep[e.g.][]{Hines2013,Riva2009}. Additionally, it is necessary to make simplifying assumptions about the nature of afterslip. For example, one can assume a frictional model for afterslip and parameterize afterslip in terms its the unkown rheologic properties \citep[e.g.][]{Johnson2009,Johnson2004}. Additionally, one can assume that afterslip only dominates the first year of deformation \citep[e.g.][]{Pollitz2012,Spinler2015} and that viscoelastic relaxation is the dominant mechanism in later years. However, numerous observations of interseismic fault creep has been recognized in this and similar tectonic settings and it has been speculated that such creep can be initiated as a postseismic process \citep{Cakir2012,Cetin2014}.  We therefore do not make to implicit assumption that afterslip terminates after a given amount of time.  Indeed, the preferred viscoelastic model from \citet{Pollitz2012} underestimates near field velocities, which could be indicative of unmodeled continued afterslip.

All of the aforementioned studies model displacements observered at stations within 200 km of the El Mayor-Cucapah epicenter, while postseismic deformation in this region following previous earthquakes of similar magnitude has been observed at distances extended out to about 300 km \citep{Freed2007a}. In the present study we examine stations within 400 km of the El Mayor-Cucapah epicenter, which turns out to be crucial for discerning the postseismic deformation mechanism.

Clearly, both afterslip and viscoelastic relaxation are involved in postseismic deformation neglecting to model one of these deformation mechanisms could result in a biased estimate of the other.  In this paper we use the inverse method described in \citet{Hines2015} to estimate the afterslip an effective viscosity necessary to describe the transience postseismic deformation over the first ten months after the El Mayor-Cucapah earthquake. We then form a suite of models which have various lithospheric rheologies but viscosities that are consistent with the effective viscosity estimated from the early postseismic deformation. Of the suite of models tested, we find that five years of postseismic deformation can be explained by a combination of sustained afterslip on the coseismicly ruptured fault and a Zener rheology upper mantle with viscosity that decays from $\sim5\times10^{18}$ to $\sim1\times10^{18}$ Pa s.

@@

We use continuous GPS position time series provided by University Navstar Consortium (UNAVCO) stations within a 400 km radius about the El Mayor-Cucapah epicenter. Our analysis is on the coseismic and postseismic deformation resulting from the El Mayor-Cucapah earthquake, which we collectively describe as $u(t)$. We consider GPS position time series, $u_\mathrm{obs}(t)$, to be the superposition of $u(t)$, secular tectonic deformation, annual and semi-annual fluctuations, and coseismic offsets from significant earthquakes over the time span of this study.  The June 14, 2010 Mw5.8 Ocotillo earthquake and the August 26, 2012 Brawley swarm, which consisted of a Mw5.5 and Mw5.3 event, are the only earthquakes after the EMC earthquake that produced noticeable offsets recorded by GPS. Although the Ocotillo earthquake had its own series of aftershocks (Haukson), neither earthquake produced transient deformation that is detectable with GPS. We thus model $u_\mathrm{obs}(t)$ as

@@

Stations which recorded signals that clearly cannot be described by the aforementioned processes are not included in our analysis. This includes stations in the Los Angeles basin, which record deformation that is largely anthropogenic. In order to ensure an accurate estimation of the secular deformation, we only use stations that were installed at least six months prior to El Mayor-Cucapah earthquake. While several stations were installed after the EMC earthquake to improve the spatial resolution of postseismic deformation \citep{Spinler2015}, our inverse method uses postseismic displacements rather than velocities (e.g. ...)
ELABORATE ON VELOCITY VS DISPLACEMENT
, which requires the knowledge of the stations preseismic position. Despite our inability to utilize potentially rich data, we prefer to use postseismic displacements rather than potentially dubious estimates of postseismic velocities.

The October 16, 1999 Hector Mine earthquake, which occurred within our study region about 270 km north of the EMC epicenter, has produce transient postseismic deformation which we do not wish to model either mechanically or through empirical line fitting. We thus restrict our analysis to deformation observed six years after the Hector Mine earthquake, past which point postseismic deformation for nearfield sites occurs at an approximately steady rate \citep{Savage2009}. When considering stations further away from the Hector Mine epicenter, postseismic transience persists for only about two years \citep{Spinler2015}.

Studies of postseismic deformation typically assume a parametric form for $u(t)$ such as such as a logarithmic or exponential function. Such functions have been found empirically e.g. \citep{Savage2005a} and theoretically \citep{Marone1991} to adequately describe deformation resulting from postseismic deformation. However, by assuming a logarithmic or exponential form of $u(t)$ we run the risk of overfitting the GPS time series and inferring a nonexistent postseismic signal. We therefore do not assume any parameteric form for $u(t)$ and rather treat it as integrated Brownian motion, so that 
\begin{equation}
    \dot{u}(t) = \sigma^2\int_0^t w(s) ds.
\end{equation}

@@

The smoothness of $u(t)$ is controlled by the chosen value of $\sigma^2$, which describes how rapidly we expect the postseismic signal to vary over time.  Setting $\sigma^2$ equal to zero will effectively result in modeling $u(t)$ as a straight line which is insufficient to describe the expected transient behavior in postseismic deformation \citep{Savage2005a}. The other endmember, where $\sigma^2$ is infinitely large, will result in $u_\mathrm{pred}(t)$ fitting what is obviously noise in the data. While one can use a maximum likelihood based approach to picking $\sigma^2$ (e.g. \citep{Segall1997}), we rather take a subjective approach and choose a value for $\sigma^2$, which is used when filtering time series for each station, that is just large enough to faithfully describes the early rapid rate of postseismic deformation at the most near field station in our study, P496.  This ensures that $\sigma^2$ will be sufficiently large so that our estimate of $u(t)$ does not smooth out potentially valuable postseismic signal. We find that when using $\sigma^2 = 0.05 \mathrm{m}^2 / \mathrm{yr}^3$, we are able to adequately describe all but the first week of postseismic deformation at station P496, which gets incorporated into our estimate of coseismic displacements (figure \ref{P496Fit}).  Since we assume that the first week of deformation is overwhelmingly the result of afterslip, any unmodeled afterslip over the first week following the El Mayor-Cucapah earthquake will be added into our estimates of coseismic slip in section ().  Figure \ref{P496PS} shows the estimate of coseismic and postseismic deformation, $u(t)$, for station P496 along with its formal uncertainties.

It is important to note that the shown uncertainties in $u(t)$ do not account for the non-negligible epistemic uncertainty in eq. \ref{TimeSeriesModel}.  For example, we assume that the background rate of deformation can be approximated as having a constant rate, while mechanical models would suggest this to not be true on earthquake cycle timescales \citep{Thatcher1983}. Although based on inspection,  assuming a constant rate of background deformation appears to be appropriate for all stations except for perhaps the stations closest to the Hector Mine epicenter, where postseismic deformation persists.  Also, our model for seasonal deformation in eq. \ref{TimeSeriesModel} assumes a constant amplitude over time, which means that years of particularly heavy or light rainfall cannot be adequately described.  This deficiency in our seasonal model causes our estimate of $u(t)$ to describe some of the unmodeled oscillations (figure \ref{P619PS}).

@@

Another reason for constraining slip to be shallower than 15 km depth is because we seek to estimate a lower crustal viscosity and the inverse problem becomes particularly ill-posed when trying to simultaneously estimate afterslip and a lithospheric viscosity at the same depths. 
This ill-posedness is illustrated in figure (\ref{fig:LowerCrust}), which shows the displacement resulting from a meter of slip on a fault extending from 15 to 30 km depth as well as the initial velocity resulting from viscoelastic relaxation in the lower crust, which is given a viscosity of $10^{18}$ Pa s.  The horizontal displacement from fault slip is in the opposite direction as the displacements resulting from subsequent viscoelastic relaxation.  This means that displacement resulting from afterslip at lower crustal depths can be cancelled out by a low viscosity lower crust.  We eliminate this null space by allowing only one mechanism in the lower crust, which we choose to be viscoelastic relaxation.  Aside from the issue of nonuniqueness, it is also nonsensical to allow for both brittle and ductile deformation mechanisms at the same depth.

@@

Most postseismic studies assume Maxwell viscoelasticity in the lower crust and upper mantle, which is the simplest viscoelastic rheologic model (e.g. \cite{Nur1974}, \cite{Pollitz2000}, \cite{Hetland2003},\cite{Freed2006a}, \cite{Johnson2009}, \cite{Hearn2009}).  In Southern California, postseismic studies following the Landers \cite{Pollitz2000}, Hector Mine \cite{Pollitz2001}, and El Mayor-Cucuapah earthquake \cite{Spinler2015}, \cite{Rollins2015}, have assumed Maxwell viscoelasticity in the lithosphere and asthenosphere and have inferred upper mantle viscosities ranging from $10^{17}$ to $10^{18}$ Pa s with a relatively stronger lower crust. While these inferences are consistent with our effective viscosities, they are inconsistent with viscosity estimates made from geophysical processes that occur over longer time scales. Studies on interseismic deformation \cite{Lundgren2009} and lake loading \cite{Luttrell2007}, which are both processes that occur on the timescale of $~10^2$ years, infer a Maxwell viscoelastic upper mantle with a viscosity on the order of $10^{19}$ Pa s.

An additional deficiency in the Maxwell rheology is that it predicts a steady decay over time in the rate of postseismic deformation, which fails to describe the commonly observed rapid early transience followed by a relatively steady rate postseismic deformation.  Early transient postseismic deformation can be attributed to fault creep \cite{Savage2005a} and the trend in postseismic deformation can then be explained with a combination of fault creep and relaxation in a Maxwell viscoelastic lithosphere, (e.g. \cite{Hearn2009}, \cite{Johnson2009}). However, the trend in postseismic deformation at distances greater than about 200 km from the El Mayor Cucapah epicenter, which can only be attributed to viscoelastic relaxation \cite{Freed2007a}, cannot be explained with a Maxwell viscoelastic lithosphere (figure \ref{fig:FinalFarFieldRS}).

We found that a Zener rheology in the upper mantle with a transient viscosity of about $10^{18}$ Pa s does a noticeably better job at predicting far field postseismic deformation.  A generalization of the Zener viscoelastic model, consisting of several Kelvin elements connected in series, is commonly used to describe the spectrum of frequencies where seismic attenuation is observed \cite{Liu1976}.  The highest viscosity needed to describe attenuation of earth's lowest frequency normal modes is on the order of $10^{16}$ Pa s \cite{Yuen1982} and the largest characteristic relaxation time would be on the order of days. Even though our inferred transient viscosity is orders of magnitude larger than that required for seismic attenuation models, the two models are not incompatible.  Rather the delayed elasticity in seismic attenuation models occurs on such short timescales that it can be considered  part of the instantaneous elastic phase of deformation associated with the preferred Zener model in this study.

Of course, it has long been recognized that a Zener rheology provides an incomplete descriptions of the asthenosphere, as it does not have the fluid behaviour required for mantle dynamics on geologic time scales. \cite{Yuen1982} proposed a Burgers rheology with a low viscosity ($\approx 10^{16}$ Pa s) in the Kelvin element and high viscosity ($\approx 10^{21}$ Pa s) in the Maxwell element which is capable of describing both seismic attenuation and the fluid-like behaviour required for glacial isostatic rebound and mantle convection. The justification of a Burger's rheology mantle is further supported by laboratory experments on the strength of olivine \cite{Chopra1997}. \cite{Pollitz2003} sought to describe postseismic deformation following Hector Mine with a Burgers rheology mantle and they found a best fitting transient viscosity of $1.6\times10^{17}$ Pa s and steady state viscosity of $4.6\times10^{18}$ Pa s. While the Burgers rheology was introduced as a means of bridging to gap between relaxation observed in long and short term geophysical processes, the inferred steady state viscosity from \cite{Pollitz2003} is still inconsistent with the Maxwell viscosities inferred from \cite{Lundgren2009} and \cite{Luttrell2007}. Additionally, the transient viscosity inferred by \cite{Pollitz2003} is needed to describe the earliest phase of postseismic deformation following the Hector Mine earthquake, which may be the result of shallow afterslip. While \cite{Pollitz2003} ruled out deep afterslip as an alternative mechanism based on inconsistent vertical deformation, it is still possible to successfully describe all components of early postseismic deformation following the Hector Mine earthquake with afterslip at seismogenic depths \cite{Jacobs2002}. It is then possible that the preferred rheologic model from \cite{Pollitz2003} was biased towards inferring a particularly low transient viscosity by neglecting to account for afterslip.  This is in contrast to the present study, where we have inferred a lithospheric viscosity structure simultaneously with afterslip. While we also argue that a transient rheology is necessary to explain postseismic deformation, our preferred transient viscosity of $10^{18}$ Pa s is an order of magnitude larger than the transient viscosity found by \cite{Pollitz2003}.  Since we found that A Zener model is able to describe the available postseismic deformation following the El Mayor-Cucapah earthquake, any Burgers rheology with a steady state viscosity that is effectively infinite over five years, would also be able to describe postseismic deformation. Such a Burgers model would then be consistent with the steady state viscosities necessary for lake loading, interseismic deformation, and mantle dynamics.

@@

\begin{figure}
\includegraphics[scale=0.9]{Figures/rheology}
\centering 
\caption{}
\label{fig:Rheology}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.9]{Figures/near_field_record_section}
\centering 
\caption{near field record section}
\label{fig:NearFieldRS}
\end{figure}
\begin{figure}
\includegraphics[scale=0.9]{Figures/far_field_record_section}
\centering 
\caption{far field record section}
\label{fig:FarFieldRS}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{Figures/lower_crust}
\centering 
\caption{afterslip and viscoelastic relaxation in lower crust}
\label{fig:LowerCrust}
\end{figure}

\begin{figure}
\includegraphics[scale=1.0]{Figures/slip_lcurve}
\centering 
\caption{near field record section}
\label{fig:SlipLCurve}
\end{figure}
